---
title: "Plethodon_Hydroclimate_Size_at_Maturity"
author: "N/A"
date: "2026-09-21"
output: html_document
---

Loading libraries (Uncomment to install)
```{r, message = FALSE}
# install.packages(c("tidyverse", "ggeffects", "patchwork"))
library(tidyverse)
library(ggeffects)
library(patchwork)
```

Reading files
```{r}
Plethodon_data <- read.csv('../Input/Plethodon_Data.csv')
Plethodon_Study_Sites <- read.csv('../Input/Salamander_Study_Sites.csv')
```

Filtering to mature Plethodon
```{r}
# Keeps mature males and females with SVL from 40 to 120 mm
all_mature_plethodon <- Plethodon_data %>%
  filter(maturity_status %in% c("mature_male", "mature_female"),
         between(svl_mm, 30, 120)) %>%
  select(site, species, svl_mm, maturity_status)
```

Minimum SVL per site and sex, combined with site environmental data
```{r}
# Smallest mature SVL per site and sex
Plethodon_sites_simplified <- all_mature_plethodon %>%
  group_by(site, maturity_status) %>%
  summarise(svl_mm = min(svl_mm), .groups = "drop") %>%
  mutate(sex = if_else(maturity_status == "mature_male", "male", "female"))

# Keeps only sites with measured mature individuals
Modeling_Data <- Plethodon_Study_Sites %>%
  inner_join(Plethodon_sites_simplified, by = "site", relationship = "one-to-many")
```

Prepping the Plethodon data for modeling
```{r}
# Standardizes each site once
scale_by_site <- function(x, site) {
  once <- !duplicated(site)
  (x - mean(x[once])) / sd(x[once])
}

Modeling_Data <- Modeling_Data %>%
  mutate(
    # Folds aspect so 0 = north-facing and 180 = south-facing
    Aspect_folded = 180 - abs(Aspect - 180),

    # Standardizes the covariates
    MDT_z    = scale_by_site(DaymetMDTC.Apr.Oct.2011.2020, site),
    MDP_z    = scale_by_site(DaymetMDPmm.Apr.Oct.2011.2020, site),
    MDVPD_z  = scale_by_site(DaymetMDVPD.Apr.Oct.2011.2020, site),
    Aspect_z = scale_by_site(Aspect_folded, site),
    Slope_z  = scale_by_site(Slope, site),

    # Sets male as the reference level
    sex = relevel(factor(sex), ref = "male"))
```

Linear model for Coweeta
```{r}
# Subsetting data to only Coweeta
Coweeta_Modeling_data <- Modeling_Data %>%
  filter(Study.Area == "Coweeta")

# Linear Model
Coweeta_Model <- glm(svl_mm ~ sex + MDVPD_z + Slope_z + Aspect_z + MDP_z + MDT_z, data = Coweeta_Modeling_data)

# Model summary
summary(Coweeta_Model)
```

Linear model for regional dataset
```{r}
# Linear Model
All_Sites_Model <- glm(svl_mm ~ sex + MDVPD_z + Slope_z + Aspect_z + MDP_z + MDT_z, data = Modeling_Data)

# Model summary
summary(All_Sites_Model)
```

Plotting
Back-transforming values for plotting
```{r}
# One row per site, matching how the covariates were standardized
Site_Data <- Modeling_Data %>%
  distinct(site, .keep_all = TRUE)

# Back-transformation values
MDVPD_mean <- mean(Site_Data$DaymetMDVPD.Apr.Oct.2011.2020)
MDVPD_sd   <- sd(Site_Data$DaymetMDVPD.Apr.Oct.2011.2020)

Slope_mean <- mean(Site_Data$Slope)
Slope_sd   <- sd(Site_Data$Slope)
```

Generating prediction objects
```{r}
# Coweeta predictions
# MDVPD
MDVPD_pred_coweeta <- ggpredict(Coweeta_Model, terms = c("MDVPD_z", "sex"))
MDVPD_pred_coweeta$MDVPD <- MDVPD_pred_coweeta$x * MDVPD_sd + MDVPD_mean

# Slope
Slope_pred_coweeta <- ggpredict(Coweeta_Model, terms = c("Slope_z", "sex"))
Slope_pred_coweeta$Slope <- Slope_pred_coweeta$x * Slope_sd + Slope_mean

# All site predictions
# MDVPD
MDVPD_pred_all_sites <- ggpredict(All_Sites_Model, terms = c("MDVPD_z", "sex"))
MDVPD_pred_all_sites$MDVPD <- MDVPD_pred_all_sites$x * MDVPD_sd + MDVPD_mean

# Slope
Slope_pred_all_sites <- ggpredict(All_Sites_Model, terms = c("Slope_z", "sex"))
Slope_pred_all_sites$Slope <- Slope_pred_all_sites$x * Slope_sd + Slope_mean
```

Calculating shared limits
```{r}
# Both MDVPD plots begin and end with the Coweeta prediction range for comparison of the Coweeta plot to the broader regional plot.
MDVPD_limits <- range(MDVPD_pred_coweeta$MDVPD)

# Both slope plots cover the combined Coweeta and all-sites range
Slope_limits <- range(Slope_pred_coweeta$Slope, Slope_pred_all_sites$Slope)

# All four plots share one SVL axis that fits every confidence interval
SVL_limits <- range(
  MDVPD_pred_coweeta$conf.low,   MDVPD_pred_coweeta$conf.high,
  Slope_pred_coweeta$conf.low,   Slope_pred_coweeta$conf.high,
  MDVPD_pred_all_sites$conf.low, MDVPD_pred_all_sites$conf.high,
  Slope_pred_all_sites$conf.low, Slope_pred_all_sites$conf.high)

# Adding some space above and below the confidence intervals
SVL_padding <- diff(SVL_limits) * 0.05
SVL_limits <- c(SVL_limits[1] - SVL_padding, SVL_limits[2] + SVL_padding)
```

Coweeta MDVPD plot
```{r}
Coweeta_MDVPD_Plot <- ggplot(MDVPD_pred_coweeta, aes(x = MDVPD, y = predicted, color = group, fill = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, color = NA) +
  geom_line(linewidth = 1.2) +
  coord_cartesian(xlim = MDVPD_limits, ylim = SVL_limits) +
  labs(title = "Coweeta", x = "MDVPD", y = "Predicted min SVL (mm)",
       color = "Sex", fill = "Sex", tag = "A") +
  theme_classic() +
  theme(legend.position = "none")
```

Coweeta slope plot
```{r}
Coweeta_Slope_Plot <- ggplot(Slope_pred_coweeta, aes(x = Slope, y = predicted, color = group, fill = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, color = NA) +
  geom_line(linewidth = 1.2) +
  coord_cartesian(xlim = Slope_limits, ylim = SVL_limits) +
  labs(title = "Coweeta", x = "Slope", y = NULL,
       color = "Sex", fill = "Sex", tag = "B") +
  theme_classic() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.position = "none")
```

All sites MDVPD plot
```{r}
All_Sites_MDVPD_Plot <- ggplot(MDVPD_pred_all_sites, aes(x = MDVPD, y = predicted, color = group, fill = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, color = NA) +
  geom_line(linewidth = 1.2) +
  coord_cartesian(xlim = MDVPD_limits, ylim = SVL_limits) +
  labs(title = "All sites", x = "MDVPD", y = "Predicted min SVL (mm)",
       color = "Sex", fill = "Sex", tag = "C") +
  theme_classic() +
  theme(legend.position = "none")
```

All sites slope plot
```{r}
All_Sites_Slope_Plot <- ggplot(Slope_pred_all_sites, aes(x = Slope, y = predicted, color = group, fill = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, color = NA) +
  geom_line(linewidth = 1.2) +
  coord_cartesian(xlim = Slope_limits, ylim = SVL_limits) +
  labs(title = "All sites", x = "Slope", y = NULL,
       color = "Sex", fill = "Sex", tag = "D") +
  theme_classic() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```

Combine all four plots into one for comparison
```{r}
# Coweeta plots on the top row, all-sites plots on the bottom row, shared legend above
Combined_Covariate_Plot <- (
  guide_area() /
    (Coweeta_MDVPD_Plot | Coweeta_Slope_Plot) /
    (All_Sites_MDVPD_Plot | All_Sites_Slope_Plot)) +
  plot_layout(heights = c(0.10, 1, 1), guides = "collect") &
  theme(legend.position = "top")

Combined_Covariate_Plot

# Saves the combined plot, creating the Outputs folder if it doesn't exist
dir.create("../Outputs", showWarnings = FALSE)
ggsave("../Outputs/combined_MDVPD_Slope_plot.pdf", plot = Combined_Covariate_Plot, width = 14, height = 9, units = "in")
```

Session information
```{r}
# Records the R and package versions used
sessionInfo()
```


